---
title: "Hockey sticks and crosses"
author: "Yusuf Imaad Khan"
format: html
editor: source
backgroundcolor: "#e5e8de"
execute:
  echo: false
  warning: false
  message: false
  #cache: true # be careful with this one
code-tools: true
toc: true
number-sections: true
toc-depth: 1
number-depth: 1
self-contained: true # deprecated in later version 
quarto:
  dependencies:
    - html2canvas@^1.3.2
---

```{r}

# Load packages

pacman::p_load(tidyverse, readr, readxl, janitor, here)

# Get handy functions

source(here("newsletter-charts", "functions", "handy-functions.R"))

# Import all CSVs from folder

csv_data <- import_csvs_from_folder("outputs")

```

```{=html}
<script src="https://cdnjs.cloudflare.com/ajax/libs/html2canvas/1.3.2/html2canvas.min.js"></script>

<!-- use relative file path -->

<script src="../functions/exportAsPNGv2.js"></script>
```

```{r}
#| cache: false

# convert to observable ----

# 
ojs_define(WID_p1vp50_US_convert = WID_p1vp50_US)

#
ojs_define(employment_by_economic_sector_convert = employment_by_economic_sector)

# 
ojs_define(tax_burden_data_convert = tax_burden_data)

# 
ojs_define(df_chn_usa_convert = df_chn_usa)

```


# Left-wing pop: cross of income

::: {#chart5 style="width: 650px; background-color: #e5e8de;"}
```{=html}
<svg id="Layer_1" data-name="Layer 1" xmlns="http://www.w3.org/2000/svg" viewBox="0 0 100 1" style="display: block; height: 100%; width: 100%; margin-top: auto; margin-left: auto"><rect x="0" y="0" width="10px" height="3px" style="fill:#000000;"/></svg>
```
::: {style="font-size: 1.3em; margin-top: -1rem; color: #000000;"}
### The cross of income shares in the US
:::

::: {style="color: #5a6570; margin-bottom: -1rem;"}
Shares of national income in the US (%) - top 1% vs bottom 50%
:::

```{ojs}

newPlot = import("https://esm.sh/@observablehq/plot");

//Parse time
parseTime = d3.utcParse("%Y");

// Transpose
data5 = transpose(WID_p1vp50_US_convert);


date5Edit = data5 // have to define this otherwise it throws an error
  .forEach((d) => {
  d.year = parseTime(d.year);
});

// Plot function - start with rough function and then add custom styling

crossLineY = (data, {x, y, z, text, stroke, color, fill,...options}) => {
  return Plot.marks(
    //Plot.ruleY([0]),
    Plot.lineY(data, {x, y, z, stroke: "white", color: "white",curve: "natural", strokeWidth: 6,...options}),
    Plot.lineY(data, {x, y, z, stroke, color, curve: "natural", strokeWidth: 4,...options}),
    Plot.text(data, Plot.selectLast({x, y, z, text, color, textAnchor: "start",fontSize:"18px",fill,dx: 10,stroke: "#e5e8de", strokeWidth: 5,...options}))
  );
}

// This seems like a poorly written function. Leave it for now - the other crosses might have differences

crossLineY(data5, {x: "year", y:"sptinc_z_us_pre_tax_national_income_top_1_percent_share_usa", z:"percentile", stroke: "percentile", text: "percentile", color: "percentile", fill: "percentile"}).plot({

marginTop:5,
marginBottom: 40,
marginLeft: 45,
marginRight: 65,

style: {fontSize: "18px", backgroundColor: "#e5e8de"},

x: {
label: null,
nice: true,
tickPadding: 10,
ticks: 5
},

y: {
label: null,
percent: true,
nice: true,
ticks: 5,
tickSize:0,
grid: true,
tickFormat: (d) => `${d}%`, 
},

color: {
legend: false,
range: ["#3f6bee", "#e61f00"]
}

})


```

::: {style="font-size: 13px; color: #5a6570;"}
**Source:** World Inequality Database <br> **Graphic:** Yusuf Imaad Khan / @yusuf_i_k
:::

::: {style="margin-top: -45px; margin-right: 10px;"}
![](../polycrisis-black.svg){height="25" fig-align="right"}
:::
:::

```{=html}
<button id="downloadButton" onclick="exportAsPNGv2('chart5', 15)">Export as PNG</button>
```
# Right-wing pop narrative: deindustrialization

::: {#chart6 style="width: 650px; background-color: #e5e8de;"}
```{=html}
<svg id="Layer_1" data-name="Layer 1" xmlns="http://www.w3.org/2000/svg" viewBox="0 0 100 1" style="display: block; height: 100%; width: 100%; margin-top: auto; margin-left: auto"><rect x="0" y="0" width="10px" height="3px" style="fill:#000000;"/></svg>
```
::: {style="font-size: 1.3em; margin-top: -1rem; color: #000000;"}
### The cross of types of work in the US
:::

::: {style="color: #5a6570; margin-bottom: -1rem;"}
US employment by economic sector - number of people (% share)
:::

```{ojs}

// Transpose
//data6v2 = transpose(shares_of_gdp_by_economic_sector_convert);

data6v2 = transpose(employment_by_economic_sector_convert);


date6v2Edit = data6v2 // have to define this otherwise it throws an error
  .forEach((d) => {
  d.year = parseTime(d.year);
});

// Plot function - start with rough function and then add custom styling

crossLineY6v2 = (data, {x, y, z, text, stroke, color, fill,...options}) => {
  return Plot.marks(
    Plot.ruleY([0]),
    Plot.lineY(data, {x, y, z,marker: "dot", stroke: "white", color: "white",curve: "natural", strokeWidth: 3,...options}),
    Plot.lineY(data, {x, y, z,marker: "dot", stroke, color, strokeWidth: 2,curve: "natural",...options}),
    Plot.text(data, Plot.selectLast({x, y, z, text, color, textAnchor: "start",fontSize:"18px",fill,dx: 10, stroke: "#e5e8de", strokeWidth: 5, lineWidth: 10,...options}))
  );
}

crossLineY6v2(data6v2, {x: "year", y:"share", z:"category", stroke: "category", text: "labels", color: "category", fill: "category"}).plot({

marginTop:20,
marginBottom: 40,
marginLeft: 50,
marginRight: 120,

style: {fontSize: "18px", backgroundColor: "#e5e8de"},

x: {
label: null,
nice: true,
tickPadding: 10,
ticks: [new Date("1840-01-01"), new Date("1900-01-01"), new Date("1950-01-01"), new Date("2015-01-01")]
},

y: {
label: null,
percent: true,
nice: false,
ticks: [0, 20, 40, 60, 80, 100],
tickSize:0,
grid: true,
tickFormat: (d) => `${d}%`, 
},

color: {
legend: false,
range: ["#e61f00","#3f6bee"]
}

})


```

::: {style="font-size: 13px; color: #5a6570;"}
**Sources:** Our World In Data; Herrendorf et al. (2014); WDI <br> **Graphic:** Yusuf Imaad Khan / @yusuf_i_k
:::

::: {style="margin-top: -45px; margin-right: 10px;"}
![](../polycrisis-black.svg){height="25" fig-align="right"}
:::
:::

```{=html}
<button id="downloadButton" onclick="exportAsPNGv2('chart6', 15)">Export as PNG</button>
```

# Cross of Tax Burdens

::: {#chart7 style="width: 650px; background-color: #e5e8de;"}
```{=html}
<svg id="Layer_1" data-name="Layer 1" xmlns="http://www.w3.org/2000/svg" viewBox="0 0 100 1" style="display: block; height: 100%; width: 100%; margin-top: auto; margin-left: auto"><rect x="0" y="0" width="10px" height="3px" style="fill:#000000;"/></svg>
```
::: {style="font-size: 1.3em; margin-top: -1rem; color: #000000;"}
### The cross of tax burdens in the US
:::

::: {style="color: #5a6570; margin-bottom: -1rem;"}
US federal corporate vs payroll taxes as a share of GDP (%)
:::

```{ojs}

// Transpose
data8 = transpose(tax_burden_data_convert);


date8Edit = data8 // have to define this otherwise it throws an error
  .forEach((d) => {
  d.year = parseTime(d.year);
});

// Plot function - start with rough function and then add custom styling

crossLineYtax = (data, {x, y, z, text, stroke, color, fill,...options}) => {
  return Plot.marks(
    Plot.ruleY([0]),
    Plot.lineY(data, {x, y, z, stroke: "white", color: "white", strokeWidth: 6,curve: "natural",...options}),
    Plot.lineY(data, {x, y, z, stroke, color, strokeWidth: 4,curve: "natural",...options}),
    Plot.text(data, Plot.selectLast({x, y, z, text, stroke: "#e5e8de", color, textAnchor: "start",fontSize:"18px",fill,dx: 10, strokeWidth: 5, lineWidth: 5,...options}))
  );
}

crossLineYtax(data8, {x: "year", y:"values", z:"categories", stroke: "categories", text: "labels", color: "categories", fill: "categories"}).plot({

marginTop:20,
marginBottom: 40,
marginLeft: 45,
marginRight: 35,

style: {fontSize: "18px", backgroundColor: "#e5e8de"},

x: {
label: null,
nice: true,
tickPadding: 10,
ticks: [new Date("1950-01-01"), new Date("1960-01-01"), new Date("1970-01-01"), new Date("1980-01-01"), new Date("1990-01-01"), new Date("2000-01-01"), new Date("2010-01-01"), new Date("2020-01-01")]
},

y: {
label: null,
nice: true,
ticks: [0, 1, 2, 3, 4, 5, 6, 7],
tickSize:0,
grid: true,
tickFormat: (d) => `${d}%`, 
},

color: {
legend: false,
range: ["#e61f00","#3f6bee"]
}

})


```

::: {style="font-size: 13px; color: #5a6570;"}
**Sources:** Office of Management and Budget; Zucman and Saez (2019) <br> **Graphic:** Yusuf Imaad Khan / @yusuf_i_k
:::

::: {style="margin-top: -45px; margin-right: 10px;"}
![](../polycrisis-black.svg){height="25" fig-align="right"}
:::
:::

```{=html}
<button id="downloadButton" onclick="exportAsPNGv2('chart7', 15)">Export as PNG</button>
```
# Additional chart: GDP

::: {#chart8 style="width: 650px; background-color: #e5e8de;"}
```{=html}
<svg id="Layer_1" data-name="Layer 1" xmlns="http://www.w3.org/2000/svg" viewBox="0 0 100 1" style="display: block; height: 100%; width: 100%; margin-top: auto; margin-left: auto"><rect x="0" y="0" width="10px" height="3px" style="fill:#000000;"/></svg>
```
::: {style="font-size: 1.3em; margin-top: -1rem; color: #000000;"}
### The cross of global power
:::

::: {style="color: #5a6570; margin-bottom: -1rem;"}
Share of world GDP (%) - PPP constant 2017 international \$
:::

```{ojs}

// Transpose
data9 = transpose(df_chn_usa_convert);


date9Edit = data9 // have to define this otherwise it throws an error
  .forEach((d) => {
  d.year = parseTime(d.year);
});

// Plot function - start with rough function and then add custom styling

crossLineYgdp = (data, {x, y, z, text, stroke, color, fill,...options}) => {
  return newPlot.marks(
    newPlot.ruleY([0]),
    newPlot.lineY(data, {x, y, z, stroke: "white", color: "white", strokeWidth: 6,curve: "natural",...options}),
    newPlot.lineY(data, {x, y, z, stroke, color, strokeWidth: 4,curve: "natural",...options}),
    newPlot.text(data, Plot.selectLast({x, y, z, text, strokeWidth: 5, stroke: "#e5e8de", color, textAnchor: "start",fontSize:"18px",lineWidth: 5, dx: 15, fill,...options})),
    newPlot.ruleX([new Date("2001-01-01")], {stroke: "black", strokeDasharray: "5,3", strokeOpacity: 0.8}),
    newPlot.axisX({
    fontWeight: (d) => (d.year === new Date("2001-01-01") ? "bold" : "normal"), // ...this did not work
    ticks: [new Date("1990-01-01"), new Date("1995-01-01"), new Date("2001-01-01"), new Date("2005-01-01"), new Date("2010-01-01"), new Date("2015-01-01"), new Date("2020-01-01")]
    }),
    newPlot.arrow(data9, {
      x1: [new Date("1998-01-01")],
      y1: 0.09,
      x2: [new Date("2000-09-09")],
      y2: 0.075,
      bend: true
    }),
    newPlot.text(data9, {
      x: [new Date("1993-01-01")],
      y: 0.095,
      text:[`China's accession to the WTO`], fill: "black", stroke: "#e5e8de", textAnchor: "start", strokeWidth: 5, lineWidth: 9, fontSize:"15px"})
  );
}

crossLineYgdp(data9, {x: "year", y:"share", z:"country_name", stroke: "country_name", text: "country_name", color: "country_name", fill: "country_name"}).plot({

height: 400,
marginTop:20,
marginBottom: 40,
marginLeft: 45,
marginRight: 35,

style: {fontSize: "18px", backgroundColor: "#e5e8de"},

x: {
label: null,
nice: true,
tickPadding: 10,
},

y: {
label: null,
percent: true,
nice: true,
ticks: 5,
tickSize:0,
grid: true,
tickFormat: (d) => `${d}%`, 
},

color: {
legend: false,
range: ["#e61f00","#3f6bee"]
}

})


```

::: {style="font-size: 13px; color: #5a6570;"}
**Source:** World Bank <br> **Graphic:** Yusuf Imaad Khan / @yusuf_i_k
:::

::: {style="margin-top: -45px; margin-right: 10px;"}
![](../polycrisis-black.svg){height="25" fig-align="right"}
:::
:::

```{=html}
<button id="downloadButton" onclick="exportAsPNGv2('chart8', 15)">Export as PNG</button>
```
